### 2018-10-22 submit/cover2.tex

\documentclass[a4paper,12pt]{letter}

% Some of the article customisations are relevant for this class

\name{Alex Kim} % To be used for the return address on the envelope
\signature{Alex Kim} % Goes after the closing (ie at the end of the letter, with space for a signature)
\address{1 Cyclotron Road, MS: 50R6048 \\ Berkeley, CA 94720-8164}
% Alternatively, these may be set on an individual basis within each letter environment.

\makelabels % this command prints envelope labels on the final page of the document

\begin{document}
\begin{letter}{The Editors\\AAS Publishing}

\opening{Hello.} % eg Hello.

This cover letter accompanies the resubmission 
{\bf Evidence for a Third Color  Parameter Within the Type~Ia Supernovae of the Nearby Supernova Factory} authored by myself on behalf
of the Nearby Supernova Factory Collaboration.  It presents a novel analysis, a first detection of a  color parameter within Type~Ia supernovae,
and analyzes timely data.  As such we believe that  it is appropriate for the Astrophysical Journal.

Also included is our response to the referees comments.  Changes to the text are included using the aastex ``trackchanges'' calls.

\closing{Regards,} % eg Regards,

%\cc{} % people this letter is cc-ed to
%\encl{} % list of anything enclosed
%\ps{} % any post scriptums. ``PS'' labels must be put in manually

\end{letter}
\end{document}